Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1) \note{\scriptsize Authors, please add/check middle initials and affiliations}

### 2014-04-14 WSSSPE1_report/WSSSPE1_summary_report.tex
Fix typos.

Two of the papers in this group focus on specific facets of software
design that are important factors in a project's sustainability but
are often addressed only late in the scientific software development
cycle, if at all: Krintz et al.~\cite{Krintz_WSSSPE} look at API
governance, and Heiland et al.~\cite{Heiland_WSSSPE} discuss maturity
models for software security. The other two papers discuss
implementation strategies for science from the perspective of
facilitating many or all facets of sustainanability-oriented software
design: Blanton and Lenhardt~\cite{Blanton_WSSSPE} contrast
large projects that have software infrastructure development built-in,
with cross-training domain scientist PIs in software engineering best
practices. Huang and Lapp~\cite{Huang_WSSSPE} discuss how various
specialized software engineering skills could be turned into shared
instrumentation with low barriers to access.

Krintz et al.~\cite{Krintz_WSSSPE} describe how in an era in which
computing frequently takes place in a distributed cloud, the control
over digital resources is increasingly shifting from physical
infrastructure to APIs, in particular web-service APIs. Yet, as Krintz
et al.~observe, unlike for physical IT infrastructure in data centers,
science communities have developed very little in the way of practices and
technology for API governance, referred to by Krintz et al.~as the
``combined policy, implementation, and deployment control''. Web APIs
can and do change, sometimes quite frequently. Raising the need to
port dependent applications. The effort required for porting is
notoriously difficult to estimate, making it nearly impossible for IT
organizations to assess and thus properly manage the impact of API
changes. To address this, Krintz et al.~propose a mechanism that
evaluates the porting effort between two versions of a web-service API
in a formal and automated way. To analyze the porting effort, they
divide API changes into \emph{syntactic similarity}, the changes in
inputs and outputs, and into \emph{semantic similarity}, the changes
in the API's behavior. In initial tests, their method showed good
congruence with human developers in scoring porting effort, offering
the possibility that API governance can become as solid a part of
scientific IT management as data center infrastructure management is
today.

@@

API governance and cybersecurity measures appropriate for a project's
maturity are all but two facets of sustainability-oriented software
development. Others include user-centered interface design, test
engineering, dependency management, and deployment operations. Each of
these facets require specialized skills and training in software
engineering. How can the implementation of best practices and
guidelines along many or all of these different facets be facilitated
in scientific software development? Blanton and
Lenhardt~\cite{Blanton_WSSSPE} contrast two models. In one, the time
and personnel devoted to software engineering is ``co-funded'' with
the driving science project. This typically implies large multi-year
collaborative projects that to succeed require significant software
infrastructure to be built, and which thus have the funding to support
one or several software engineers. The sustainability of such projects
then depends on sustaining the funding. In the other model, domain
scientists also take the role of software developers, whether by
necessity such as funding limitations, or due to cross-disciplinary
professional interests. For this to result in sustainable software
products, the domain scientists need to be (or become) cross-trained
in software engineering standards and best practices.

>>>>

Two of the papers in this group focus on specific facets of software
design that are important factors in a project's sustainability but
are often addressed only late in the scientific software development
cycle, if at all: Krintz et al.~\cite{Krintz_WSSSPE} look at API
governance, and Heiland et al.~\cite{Heiland_WSSSPE} discuss maturity
models for software security. The other two papers discuss
implementation strategies for science from the perspective of
facilitating many or all facets of sustainability-oriented software
design: Blanton and Lenhardt~\cite{Blanton_WSSSPE} contrast
large projects that have software infrastructure development built-in,
with cross-training domain scientist PIs in software engineering best
practices. Huang and Lapp~\cite{Huang_WSSSPE} discuss how various
specialized software engineering skills could be turned into shared
instrumentation with low barriers to access.

Krintz et al.~\cite{Krintz_WSSSPE} describe how in an era in which
computing frequently takes place in a distributed cloud, the control
over digital resources is increasingly shifting from physical
infrastructure to APIs, in particular web-service APIs. Yet, as Krintz
et al.~observe, unlike for physical IT infrastructure in data centers,
science communities have developed very little in the way of practices and
technology for API governance, referred to by Krintz et al.~as the
``combined policy, implementation, and deployment control''. Web APIs
can and do change, sometimes quite frequently, raising the need to
port dependent applications. The effort required for porting is
notoriously difficult to estimate, making it nearly impossible for IT
organizations to assess and thus properly manage the impact of API
changes. To address this, Krintz et al.~propose a mechanism that
evaluates the porting effort between two versions of a web-service API
in a formal and automated way. To analyze the porting effort, they
divide API changes into \emph{syntactic similarity}, the changes in
inputs and outputs, and into \emph{semantic similarity}, the changes
in the API's behavior. In initial tests, their method showed good
congruence with human developers in scoring porting effort, offering
the possibility that API governance can become as solid a part of
scientific IT management as data center infrastructure management is
today.

@@

API governance and cybersecurity measures appropriate for a project's
maturity are all but two facets of sustainability-oriented software
development. Others include user-centered interface design, test
engineering, dependency management, and deployment operations. Each of
these facets requires specialized skills and training in software
engineering. How can the implementation of best practices and
guidelines along many or all of these different facets be facilitated
in scientific software development? Blanton and
Lenhardt~\cite{Blanton_WSSSPE} contrast two models. In one, the time
and personnel devoted to software engineering is ``co-funded'' with
the driving science project. This typically implies large multi-year
collaborative projects that to succeed require significant software
infrastructure to be built, and which thus have the funding to support
one or several software engineers. The sustainability of such projects
then depends on sustaining the funding. In the other model, domain
scientists also take the role of software developers, whether by
necessity such as funding limitations, or due to cross-disciplinary
professional interests. For this to result in sustainable software
products, the domain scientists need to be (or become) cross-trained
in software engineering standards and best practices.
